---
output: pdf_document
geometry: margin=1in
---

\begin{itemize}

\item[1.] {\bf Exercise 1 -- Agresti 7.36}

Table \ref{table:exo1} is based on a study involving British doctors.

\begin{table}[h]
\begin{center}
\begin{tabular}{lrrrr}
 & \multicolumn{2}{c}{Person-Years}  &  \multicolumn{2}{c}{Coronary Deaths} \\ 
 Age & Nonsmokers & Smokers & Nonsmokers & Smokers \\ \hline
35 -- 44 & 18,793 & 52,407 & 2 & 32 \\
45 -- 54 & 10,673 & 43,248 & 12 & 104 \\
55 -- 64 & 5,710 & 28,612 & 28 & 206 \\
65 -- 74 & 2,585 & 12,663 & 28 & 186 \\
75 -- 84 & 1,462 & 5,317 & 31 & 102 \\ \hline
\end{tabular}
\caption{\bf Data on Coronary Death Rates} \label{table:exo1}
\end{center}
\end{table}

\begin{itemize}
\item[(a)] Fit a main effects model for the log rates using age and smoking as factors. In discussing lack-of-fit, show that this model assumes a constant ratio of nonsmokers' to smokers' coronary death rates over age, and evaluate how the sample ratio depends on age.

\vspace{3 in}

\item[(b)] Explain why it is sensible to add a quantitative interaction of age and smoking. For this model, show that the log ratio of coronary death rates changes linearly with age. Assign scores to age, fit the model, and interpret.
\end{itemize}
\end{itemize}







\newpage





 \begin{itemize}
     \item[2.] {\bf Exercise 2}
     
          
     \begin{itemize}
     \item[]
     One question in the 1990 General Social Survey asked subjects how many times they had sexual intercourse in the preceding month. Table \ref{table:exo2} shows responses classified by gender.
\begin{table}[h]
\begin{center}
\begin{tabular}{crr|rrr|ccc} \hline
Response & Male & Female & Response & Male & Female & Response & Male & Female \\ \hline
0 & 65 & 128 & 9 & 2 & 2 & 20 & 7 & 6 \\
1 & 11 & 17 & 10 & 24 & 13 & 22 & 0 & 1 \\
2 & 13 & 23 & 12 & 6 & 10 & 23 & 0 & 1 \\
3 & 14 & 16 & 13 & 3 & 3 & 24 & 1 & 0 \\
4 & 26 & 19 & 14 & 0 & 1 & 25 & 1 & 3 \\
5 & 13 & 17 & 15 & 3 & 10 & 27 & 0 & 1 \\
6 & 15 & 17 & 16 & 3 & 1 & 30 & 3 & 1 \\
7 & 7 & 3 & 17 & 0 & 1 & 50 & 1 & 0 \\
8 & 21 & 15 & 18 & 0 & 1 & 60 & 1 & 0 \\ \hline
\end{tabular}
\caption{Data from the 1990 General Social Survey} \label{table:exo2}
\end{center}
\end{table}

     
 \end{itemize} 
 \end{itemize}
 
  
  \ 
 
 
   
 \begin{itemize} 
     \item[(a)] Fit a Poisson GLM with log link and a dummy variable for gender (1=males, 0=females) and explain if the model seems appropriate.

 \end{itemize}
 
  \ 
  
  
```{r}
setwd("G:\\math\\661")
dat<-read.csv("sex.csv")
dat<-data.frame(
	  (rep(dat$Response,2)),
	c(dat$Male,dat$Female),
	as.factor(c(rep(1,nrow(dat)),rep(0,nrow(dat))))	)
names(dat)<-c("response","counts","gender")
str(dat)
cbind(head(dat) ,tail(dat))



dat.fit<-glm(response ~ gender, family=poisson, weights=counts, data=dat) 
summary(dat.fit)



tab<-cbind(dat[which(dat$gender == 0),],dat[which(dat$gender == 1 ),-1])
tab<-tab[,c(1,2,4)];tab[19,2]<-sum(tab[19:nrow(tab),2]);
tab[19,3]<-sum(tab[19:nrow(tab),3]);tab<-tab[1:19,]
names(tab)[2:3]<-c("Female","Male")
head(tab) 
c(sum(tab[,2]),sum(tab[,1]*tab[,2]));    sum(tab[,1]*tab[,2])/sum(tab[,2])
sum( tab[,2]*((tab[,1]- 4.183871 )^2) ) / ( sum(tab[,2]) -1)

c(sum(tab[,3]),sum(tab[,1]*tab[,3]));    sum(tab[,1]*tab[,3])/sum(tab[,3])
sum( tab[,3]*((tab[,1]- 4.183871 )^2) ) / ( sum(tab[,3]) -1)


1-pchisq(3985.7,43)
```
  
  \ 
  

 
 
 \begin{itemize}
     \item[]   The sample mean for the 1297 women is 4.183871 with a variance of 29.76867. The sample mean for the 1297 men is 5.404167 with a variance of 31.30203. In both groups the sample variances are about 6-7 times the size of the sample means. This is suggesting overdispersion relative to the Poisson. We also see that the model does not give a good fit to the data ($p$-value $\approx0$).


 \end{itemize}
 
 \ 
 
 \begin{itemize}
     \item[(b)] Interpret the regression coefficient of gender for the model in (a) and provide a 95\% Wald confidence interval for the ratio of means for males versus females.

 \end{itemize}
 
 
 
 
 
  \ 
  
 \begin{itemize}
     \item[] \vspace{1.5 in} 

 \end{itemize}  
 
 \ 
 
 
 \begin{itemize}
     \item[(c)] Fit a negative binomial model.  Is there evidence of overdispersion? What is the estimated difference in log means, its standard error, and the 95\% Wald confidence interval for the ratio of means. 

 \end{itemize}
 
 
  \ 
  
  
  
```{r}
library(MASS)
nb.fit<-glm.nb(response ~ gender, weights=counts, data=dat)
summary(nb.fit)

UGdat<-as.data.frame(lapply(dat, function(x,p) rep(x,p), dat[["counts"]]))
hist(UGdat[,1], breaks = seq(0,60,by=1))
```
  
 
 
  \ 
  
 \begin{itemize}
     \item[] \vspace{1.5 in} 

 \end{itemize}  
 
 \ 
 
 
 
 \begin{itemize}
     \item[(d)] Consider a zero-inflated Poisson model with the zero-inflated component constant across subject (that is with intercept only for the model of $\phi_i$). What are the mixing proportions for the degenerate distribution and the Poisson model? Interpret the regression coefficient of gender.

 \end{itemize}
 
 
   \ 
  

  
```{r}
suppressWarnings(suppressMessages(library(pscl)))

fit.zip = zeroinfl(response ~ gender |  1 ,data=UGdat)
summary(fit.zip )

#mixing proportions
as.numeric( exp(coef(fit.zip)[3])/(1+exp(coef(fit.zip)[3])) )
1-as.numeric( exp(coef(fit.zip)[3])/(1+exp(coef(fit.zip)[3])) )
```  
  
  
 

 
  \ 
  
 \begin{itemize}
     \item[] \vspace{1.5 in} 

 \end{itemize}  
 
 \ 
 
 
 
 \begin{itemize}
     \item[(e)] Consider a zero-inflated negative binomial model. What are the mixing proportions for the degenerate distribution and the negative binomial model? Interpret the regression coefficient of gender.

 \end{itemize}
 
 
  \ 
  
  
```{r}
fit.zinb = zeroinfl(response ~ gender| 1 ,dist="negbin",data=UGdat)
summary(fit.zinb)

#mixing proportions
as.numeric( exp(coef(fit.zinb)[3])/(1+exp(coef(fit.zinb)[3])) )
1-as.numeric( exp(coef(fit.zinb)[3])/(1+exp(coef(fit.zinb)[3])) )
```
  
  
  
  
  
 
 
  \ 
  
 \begin{itemize}
     \item[] \vspace{1.5 in} 

 \end{itemize}  
 
 \ 
 
 
 
 \begin{itemize}
\item[(f)] Provide a table with the observed counts and the fitted counts for each of the four models for $y_i=0, \ldots,20$ and $y_i > 20$.

 \end{itemize}
 
  
 
 