---
output: pdf_document
---


 \begin{itemize}
     \item[2.] {\bf Exercise 2}
     
          
     \begin{itemize}
     \item[]
     One question in the 1990 General Social Survey asked subjects how many times they had sexual intercourse in the preceding month. Table \ref{table:exo2} shows responses classified by gender.
\begin{table}[h]
\begin{center}
\begin{tabular}{crr|rrr|ccc} \hline
Response & Male & Female & Response & Male & Female & Response & Male & Female \\ \hline
0 & 65 & 128 & 9 & 2 & 2 & 20 & 7 & 6 \\
1 & 11 & 17 & 10 & 24 & 13 & 22 & 0 & 1 \\
2 & 13 & 23 & 12 & 6 & 10 & 23 & 0 & 1 \\
3 & 14 & 16 & 13 & 3 & 3 & 24 & 1 & 0 \\
4 & 26 & 19 & 14 & 0 & 1 & 25 & 1 & 3 \\
5 & 13 & 17 & 15 & 3 & 10 & 27 & 0 & 1 \\
6 & 15 & 17 & 16 & 3 & 1 & 30 & 3 & 1 \\
7 & 7 & 3 & 17 & 0 & 1 & 50 & 1 & 0 \\
8 & 21 & 15 & 18 & 0 & 1 & 60 & 1 & 0 \\ \hline
\end{tabular}
\caption{Data from the 1990 General Social Survey} \label{table:exo2}
\end{center}
\end{table}

     
 \end{itemize} 
 \end{itemize}
 
  
  \ 
 
 
   
 \begin{itemize} 
     \item[(a)] Fit a Poisson GLM with log link and a dummy variable for gender (1=males, 0=females) and explain if the model seems appropriate.

 \end{itemize}
 
  \ 
  
  
```{r}
setwd("G:\\math\\661")
dat<-read.csv("sex.csv")
dat<-data.frame(
	  (rep(dat$Response,2)),
	c(dat$Male,dat$Female),
	as.factor(c(rep(1,nrow(dat)),rep(0,nrow(dat))))	)
names(dat)<-c("response","counts","gender")
str(dat)
cbind(head(dat) ,tail(dat))



dat.fit<-glm(response ~ gender, family=poisson, weights=counts, data=dat) 
summary(dat.fit)



tab<-cbind(dat[which(dat$gender == 0),],dat[which(dat$gender == 1 ),-1])
tab<-tab[,c(1,2,4)];tab[19,2]<-sum(tab[19:nrow(tab),2]);
tab[19,3]<-sum(tab[19:nrow(tab),3]);tab<-tab[1:19,]
names(tab)[2:3]<-c("Female","Male")
tab 
c(sum(tab[,2]),sum(tab[,1]*tab[,2]));    sum(tab[,1]*tab[,2])/sum(tab[,2])
sum( tab[,2]*((tab[,1]- 4.183871 )^2) ) / ( sum(tab[,2]) -1)

c(sum(tab[,3]),sum(tab[,1]*tab[,3]));    sum(tab[,1]*tab[,3])/sum(tab[,3])
sum( tab[,3]*((tab[,1]- 4.183871 )^2) ) / ( sum(tab[,3]) -1)


1-pchisq(3985.7,43)
```
  
  \ 
  

 
 
 \begin{itemize}
     \item[]   The sample mean for the 1297 women is 4.183871 with a variance of 29.76867. The sample mean for the 1297 men is 5.404167 with a variance of 31.30203. In both groups the sample variances are about 6-7 times the size of the sample means. This is suggesting overdispersion relative to the Poisson. We also see that the model does not give a good fit to the data ($p$-value $\approx0$).


 \end{itemize}
 
 \ 
 
 \begin{itemize}
     \item[(b)] Interpret the regression coefficient of gender for the model in (a) and provide a 95\% Wald confidence interval for the ratio of means for males versus females.

 \end{itemize}
 
 
 
 
 
  \ 
  
 \begin{itemize}
     \item[] ssssssssssssssss 

 \end{itemize}  
 
 \ 
 
 
 \begin{itemize}
     \item[(c)] Fit a negative binomial model.  Is there evidence of overdispersion? What is the estimated difference in log means, its standard error, and the 95\% Wald confidence interval for the ratio of means. 

 \end{itemize}
 
 
 
 
  \ 
  
  
  
```{r}
library(MASS)
nb.fit<-glm.nb(response ~ gender, weights=counts, data=dat)
summary(nb.fit)

UGdat<-as.data.frame(lapply(dat, function(x,p) rep(x,p), dat[["counts"]]))
hist(UGdat[,1], breaks = seq(0,60,by=1))
```
  
  
 
 \ 
 
 
 \begin{itemize}
     \item[(d)] Consider a zero-inflated Poisson model with the zero-inflated component constant across subject (that is with intercept only for the model of $\phi_i$). What are the mixing proportions for the degenerate distribution and the Poisson model? Interpret the regression coefficient of gender.

 \end{itemize}
 
 
 
 
 
  \ 
  
  
  
  
 
 \ 
 
 
 \begin{itemize}
     \item[(e)] Consider a zero-inflated negative binomial model. What are the mixing proportions for the degenerate distribution and the negative binomial model? Interpret the regression coefficient of gender.

 \end{itemize}
 
 
  \ 
  
  
  
  
 
 \ 
 
 
 \begin{itemize}
\item[(f)] Provide a table with the observed counts and the fitted counts for each of the four models for $y_i=0, \ldots,20$ and $y_i > 20$.

 \end{itemize}
 
 
  \ 
  
  
  
  
 
 \ 
 
 
 