---
output: pdf_document
geometry: margin=1in
header-includes:
  - \usepackage{bm}
---

```{r set-options, echo=FALSE, cache=FALSE}
options(width = 1111)
```

\begin{itemize}

\item[1.] {\bf Exercise 1 -- Agresti 7.36}

Table \ref{table:exo1} is based on a study involving British doctors.

\begin{table}[h]
\begin{center}
\begin{tabular}{lrrrr}
 & \multicolumn{2}{c}{Person-Years}  &  \multicolumn{2}{c}{Coronary Deaths} \\ 
 Age & Nonsmokers & Smokers & Nonsmokers & Smokers \\ \hline
35 -- 44 & 18,793 & 52,407 & 2 & 32 \\
45 -- 54 & 10,673 & 43,248 & 12 & 104 \\
55 -- 64 & 5,710 & 28,612 & 28 & 206 \\
65 -- 74 & 2,585 & 12,663 & 28 & 186 \\
75 -- 84 & 1,462 & 5,317 & 31 & 102 \\ \hline
\end{tabular}
\caption{\bf Data on Coronary Death Rates} \label{table:exo1}
\end{center}
\end{table}

\begin{itemize}
\item[(a)] Fit a main effects model for the log rates using age and smoking as factors. In discussing lack-of-fit, show that this model assumes a constant ratio of nonsmokers' to smokers' coronary death rates over age, and evaluate how the sample ratio depends on age.
\end{itemize}
\end{itemize}


\ 

```{r}
smokers<-data.frame(	c(  35 , 44 , 18793 , 52407 , 2 , 32 ),	
  c( 45 , 54 , 10673 , 43248 , 12 , 104 ),	c( 55 , 64 , 5710 , 28612 , 28 , 206 ),	
  c( 65 , 74 , 2585 , 12663 , 28 , 186 ),	c( 75 , 84 , 1462 , 5317 , 31 , 102 )  )

smokers<-t(as.matrix(smokers));smokers<-as.data.frame(smokers)
rownames(smokers)<-paste0( smokers$V1,"-", smokers$V2); smokers<-smokers[,-c(1:2)]
names(smokers)<-c(paste0("PY",c("nonsmokers", "smokers")),
	paste0("CD",c("nonsmokers", "smokers")))

smokers<-data.frame(rep(1:5,2),c(rep("NS",5),rep("S",5))	,
	c(smokers$PYnonsmokers,smokers$PYsmokers),
	c(smokers$CDnonsmokers,smokers$CDsmokers))

names(smokers)<-c("age","smoker","PersonYears","Deaths")
smokers$age<-as.factor(smokers$age);smokers$ageQI<-as.numeric(smokers$age)
smokers$ratios<-smokers[,4]/smokers[,3];str(smokers)

smokers.fit<- glm(Deaths ~ age+smoker, offset = log(PersonYears), family=poisson, data=smokers) 
summary(smokers.fit)

par(mar=c(5.1,4.1,2.1,2.1))
plot( ( smokers.fit$linear.predictors[1:5] )  - log(smokers$PersonYears)[1:5],
	 ylim=c(-8,-3.75) , pch=22 ,col="blue", bg ="blue", 
		ylab="log(death rates)",xlab="age")  
points(smokers.fit$linear.predictors[6:10]  - log(smokers$PersonYears)[6:10],
	pch=21 , bg ="red",col="red")
legend(3.5,-7,c("Non-Smokers", "Smokers"), pch = c(22,21), col=c("blue","red") , pt.bg=c("blue","red"))
```


\ 



\begin{itemize} \item[] \begin{itemize} \item[] 
We note the rate of death is modeled by:

\hfil \(\log\left( \cfrac{\mu_i}{t_i}\right) = {\bm x^T_i} {\bm \beta} \).

R is using a nequivalent offset for calculating the linear predictors,

\ 

\hfil \(  \log(\mu_i)  = \log(t_i) + {\bm x^T_i} {\bm \beta}  \).

\ 

So we move the $\log(t_i)$ term back to the LHS to show the constant ratio of coronary deaths between nonsmokers to smokers. (Note: even if we kept the $\log(t_i)$ term on the RHS, the model would still have a constant ratio between nonsmokers to smokers coronary death counts).

\ 

We see the lack-of-fit with the model:

\end{itemize} \end{itemize}

\ 

```{r}
1-pchisq(12.134  , 4 )
```
 


```{r, fig1, fig.height = 4 , fig.width = 6.5}
par(mar=c(5.1,4.1,1.1,2.1))
plot(1:5,smokers$ratios[1:5] , pch=22 ,col="blue", bg ="blue", ylab="sample ratio",xlab="age")  
points(1:5,smokers$ratios[6:10],pch=21 , bg ="red",col="red")
legend(3.75,0.005,c("Non-Smokers", "Smokers"), pch = c(22,21), col=c("blue","red") , pt.bg=c("blue","red"))
```


\vspace{3 in}

\begin{itemize} \item[] \begin{itemize}
\item[(b)] Explain why it is sensible to add a quantitative interaction of age and smoking. For this model, show that the log ratio of coronary death rates changes linearly with age. Assign scores to age, fit the model, and interpret.
\end{itemize}
\end{itemize}

```{r}
smokersQI.fit<-  glm(Deaths ~ ageQI*smoker, offset = log(PersonYears),family=poisson, data=smokers) 
summary(smokersQI.fit)
plot(1:5,-8.86716 + 1.04685*(1:5) , ylim=c(-8,-2), pch=22 ,col="blue", bg ="blue")
points(1:5,-8.86716 + 1.04685*(1:5) +1.284  -0.249,pch=21 , bg ="red",col="red")
```






\newpage





 \begin{itemize}
     \item[2.] {\bf Exercise 2}
     
          
     \begin{itemize}
     \item[]
     One question in the 1990 General Social Survey asked subjects how many times they had sexual intercourse in the preceding month. Table \ref{table:exo2} shows responses classified by gender.
\begin{table}[h]
\begin{center}
\begin{tabular}{crr|rrr|ccc} \hline
Response & Male & Female & Response & Male & Female & Response & Male & Female \\ \hline
0 & 65 & 128 & 9 & 2 & 2 & 20 & 7 & 6 \\
1 & 11 & 17 & 10 & 24 & 13 & 22 & 0 & 1 \\
2 & 13 & 23 & 12 & 6 & 10 & 23 & 0 & 1 \\
3 & 14 & 16 & 13 & 3 & 3 & 24 & 1 & 0 \\
4 & 26 & 19 & 14 & 0 & 1 & 25 & 1 & 3 \\
5 & 13 & 17 & 15 & 3 & 10 & 27 & 0 & 1 \\
6 & 15 & 17 & 16 & 3 & 1 & 30 & 3 & 1 \\
7 & 7 & 3 & 17 & 0 & 1 & 50 & 1 & 0 \\
8 & 21 & 15 & 18 & 0 & 1 & 60 & 1 & 0 \\ \hline
\end{tabular}
\caption{Data from the 1990 General Social Survey} \label{table:exo2}
\end{center}
\end{table}

     
 \end{itemize} 
 \end{itemize}
 
  
  \ 
 
 
   
 \begin{itemize} 
     \item[(a)] Fit a Poisson GLM with log link and a dummy variable for gender (1=males, 0=females) and explain if the model seems appropriate.

 \end{itemize}
 
  \ 
  
  
```{r}
setwd("G:\\math\\661")
dat<-read.csv("sex.csv")
dat<-data.frame(
	  (rep(dat$Response,2)),
	c(dat$Male,dat$Female),
	as.factor(c(rep(1,nrow(dat)),rep(0,nrow(dat))))	)
names(dat)<-c("response","counts","gender")
str(dat)
cbind(head(dat) ,tail(dat))



dat.fit<-glm(response ~ gender, family=poisson, weights=counts, data=dat) 
summary(dat.fit)

 
tab<-cbind(dat[which(dat$gender == 0),],dat[which(dat$gender == 1 ),-1])
tab<-tab[,c(1,2,4)];tab[19,2]<-sum(tab[19:nrow(tab),2]);
tab[19,3]<-sum(tab[19:nrow(tab),3]);tab<-tab[1:19,]
names(tab)[2:3]<-c("Female","Male")
head(tab) 
c(sum(tab[,2]),sum(tab[,1]*tab[,2]));    sum(tab[,1]*tab[,2])/sum(tab[,2])
sum( tab[,2]*((tab[,1]- 4.183871 )^2) ) / ( sum(tab[,2]) -1)

c(sum(tab[,3]),sum(tab[,1]*tab[,3]));    sum(tab[,1]*tab[,3])/sum(tab[,3])
sum( tab[,3]*((tab[,1]- 4.183871 )^2) ) / ( sum(tab[,3]) -1)


1-pchisq(3985.7,43)
```
  
  \ 
  

 
 
 \begin{itemize}
     \item[]   The sample mean for the 1297 women is 4.183871 with a variance of 29.76867. The sample mean for the 1297 men is 5.404167 with a variance of 31.30203. In both groups the sample variances are about 6-7 times the size of the sample means. This is suggesting overdispersion relative to the Poisson. We also see that the model does not give a good fit to the data ($p$-value $\approx0$).


 \end{itemize}
 
 \ 
 
 \begin{itemize}
     \item[(b)] Interpret the regression coefficient of gender for the model in (a) and provide a 95\% Wald confidence interval for the ratio of means for males versus females.
     
     \ 

 \end{itemize}
 
```{r}
exp(0.30850 -1.96*0.03822);exp(0.30850 +1.96*0.03822)
```
  
  \ 
  
 \begin{itemize}
     \item[] When gender is male, the count of sexual intercourse is estimated to be 1.36 times that of females ($e^{0.30850}$). The Wald 95\% confidence interval for the ratio of means for males versus females is:
     
     \hfil \(\displaystyle \exp\left( 0.30850 \pm 1.96 \cdot 0.03822 \right) = (1.263125,1.467281) \)

 \end{itemize}  
 
 \ 
 
 
 \begin{itemize}
     \item[(c)] Fit a negative binomial model.  Is there evidence of overdispersion? What is the estimated difference in log means, its standard error, and the 95\% Wald confidence interval for the ratio of means. 

 \end{itemize}
 
 
  \ 
  
  
  
```{r}
library(MASS)
nb.fit<-glm.nb(response ~ gender, weights=counts, data=dat)
summary(nb.fit)
1-pchisq(600,43)
```
  
 
 
  \ 
  
 \begin{itemize}
     \item[] We note that $\widehat{Var}(Y) = \hat{\mu} + \hat{\gamma}\hat{\mu}^2$ is actually overestimating the sample variances. For females the sample variance is 29.76867, and the negative binomial model is estimating $4.303205 + \left(\cfrac{1}{0.5019}\right) * 4.303205^2 = 41.19815$. Likewise for males, the sample variance is  31.30203, and the negative binomial model is estimating $5.858303  + \left(\cfrac{1}{0.5019}\right) * 5.858303^2=74.23789$.
     
     \ 

There is evidence that $\hat{\gamma} > 0$; $\hat{\gamma}=\left(\cfrac{1}{0.5019}  \right)=1.992429$ and a 95\% confidence interval for $\gamma$ is given by:

\hfil \( \displaystyle \cfrac{1}{0.5019 \pm 1.96 \cdot 0.0387} = \left( 1.730846 , 2.347153\right) \).

\ 

So the extra parameter is picking up some of the dispersion compared with the Poisson. But we recall that the negative binomial approaches the Poisson as $\gamma  \rightarrow 0$, so there might be more overdispersion unaccounted for. We test the fit of the negative binomial model: 
 \end{itemize}  

```{r}
1-pchisq(600.60,43)
```

 \begin{itemize}
     \item[] 
and see the model does not fit well. We look at a histogram of the raw data and see an excessive amount of zeroes in the data, one reaspon for overdispersion.
     \end{itemize}  
     
     \
     
```{r}
UGdat<-as.data.frame(lapply(dat, function(x,p) rep(x,p), dat[["counts"]]))
hist(UGdat[,1], breaks = seq(0,60,by=1))
```
 

 
 \ 
 
 
 
 \begin{itemize}
     \item[(d)] Consider a zero-inflated Poisson model with the zero-inflated component constant across subject (that is with intercept only for the model of $\phi_i$). What are the mixing proportions for the degenerate distribution and the Poisson model? Interpret the regression coefficient of gender.

 \end{itemize}
 
 
   \ 
  

  
```{r}
suppressWarnings(suppressMessages(library(pscl)))

fit.zip = zeroinfl(response ~ gender |  1 ,data=UGdat)
summary(fit.zip )

#mixing proportions
phi=as.numeric( exp(coef(fit.zip)[3])/(1+exp(coef(fit.zip)[3])) )
phi;1-phi
```  
  
  
 

 
  \ 
  
 \begin{itemize}
     \item[] The mixing parameter $\phi =  0.3505542$, so the mixing proportion for the degenerate distribution is 0.6494458 and the mixing proportion for the Poisson distribution at $y_i = 0$ is 0.3505542.
     
     \ 
     
Males have an expected log count that is 0.09242 higher than females.

 \end{itemize}  
 
 \ 
 
 
 
 \begin{itemize}
     \item[(e)] Consider a zero-inflated negative binomial model. What are the mixing proportions for the degenerate distribution and the negative binomial model? Interpret the regression coefficient of gender.

 \end{itemize}
 
 
  \ 
  
  
```{r}
fit.zinb = zeroinfl(response ~ gender| 1 ,dist="negbin",data=UGdat)
summary(fit.zinb)

#mixing proportions
phi=as.numeric( exp(coef(fit.zinb)[3])/(1+exp(coef(fit.zinb)[3])) )
phi;1-phi
```
  
  
  
  
  
 
 
  \ 
  
 \begin{itemize}
     \item[] The mixing parameter $\phi =  0.300723$, so the mixing proportion for the degenerate distribution is 0.6494458 and the mixing proportion for the Poisson distribution at $y_i = 0$ is 0.699277.
     
     \ 
     
     Males have an expected count that is 1.157011 $\left( e^{0.14584}\right)$ than females. The predictor \verb|gender1| in the part of the negative binomial regression model predicting how many times they had sexual intercourse in the preceding month is not statistically significant.

 \end{itemize}  
 
 \ 
 
 
 
 \begin{itemize}
\item[(f)] Provide a table with the observed counts and the fitted counts for each of the four models for $y_i=0, \ldots,20$ and $y_i > 20$.

 \end{itemize}
 
  \ 
  
```{r}
tab<-cbind(dat[which(dat$gender == 0),],dat[which(dat$gender == 1 ),-1])
tab<-tab[,c(1,2,5)] 
 names(tab)[2:3]<-c("Female","Male");head(tab) 

sex.n = by(dat$counts, dat$gender, sum)
sex.nb = glm.nb(response ~ gender, weights=counts, data=dat)
sex.muhat = unique(sex.nb$fitted.values)


 
nfemale.poi = dpois(unique(dat[,1]), lambda=sex.muhat[1])*sex.n[1]
nmale.poi = dpois(unique(dat[,1]), lambda=sex.muhat[2])*sex.n[2]
nfemale.negb = round(dnbinom(unique(dat[,1]), size=sex.nb$theta,
mu=sex.muhat[1])*sex.n[1],3)
nmale.negb = round(dnbinom(unique(dat[,1]), size=sex.nb$theta, 
	mu=sex.muhat[2])*sex.n[2],3)


tab$poiF<-round( nfemale.poi  ,3)
tab$poiM<-round( nmale.poi ,3)

tab$negbF<-  nfemale.negb 
tab$negbM<-  nmale.negb 


tab$zipF<- round(c(  as.numeric(
	sex.n[1] * (     phi + (1-phi)  * dpois(0,exp(1.99107)) )    ),
	sex.n[1]*( (1-phi)* dpois(unique(dat[,1])[-1],exp(1.99107)) ) ) ,3) 

tab$zipM<-round(c(  as.numeric(
	sex.n[2] * (     phi + (1-phi)  * dpois(0,exp(1.99107 + 0.09242)) )),
	sex.n[2]*((1-phi)*dpois(unique(dat[,1])[-1],exp(1.99107 + 0.09242)))),3)

tab$zinbF<-round(c( as.numeric(
	sex.n[1]*(  phi + (1-phi) * dnbinom(0,mu=exp(1.89133), exp(0.43572)))),
	sex.n[1]* ((1-phi) * dnbinom(unique(dat[,1])[-1], mu=exp(1.89133),
	exp(0.43572))   )	),3)

tab$zinbM<-round(c( as.numeric(
	sex.n[2]*(  phi + (1-phi) * dnbinom(0,mu=exp(1.89133+0.14584),
		exp(0.43572))) ),
	sex.n[2]* ((1-phi) * dnbinom(unique(dat[,1])[-1], mu=exp(1.89133 +  
		0.14584),exp(0.43572))   )),3)

print(tab,row.names = F)
```
  
 
 