---
header-includes:
  \AtBeginDocument{\let\maketitle\relax}
geometry: margin=.75in
output: pdf_document
---


```{r,echo=F}
rm(list = ls())
setwd("G:\\math\\661")
donner<-read.csv("Donner.csv",header=T)
donner<-donner[,c(2,3,4)]
donner$Male.Gender<-as.factor(donner$Male.Gender)
donner.glm<-glm(Survived~.,data=donner,family=binomial )

```

\begin{itemize}
\item[ ] 
\item[ ]  \begin{itemize}
\item[(e)]  Assess the model goodness-of-fit. 

\ 

\begin{itemize}
\item[] We conduct the Hosmer-Lemeshow goodness of fit test to assess the model goodness-of-fit. We test $H_0$: the current model fits well.


\end{itemize}


\end{itemize}
\end{itemize}


 \ 
 
```{r}
library(ResourceSelection)
res<-hoslem.test(donner.glm$y,fitted(donner.glm))
res
cbind(res$observed,res$expected)
```

\ 

\begin{itemize}
\item[ ]   
\ 

\begin{itemize}
\item[] This $p$-value of $0.1134$ is non-significant at the $0.05$ level so there should be no evidence that the model is fitting poorly. However, because the $p$-value is near the threshold there may be evidence that the model is not a great fit. With caution, we declare the model does appear adequate for these data.



\end{itemize}

 \end{itemize}

